### 2024-04-21 paper/paper.tex
added latex for the paper

\documentclass[12pt, a4paper]{article}
\usepackage[margin=5pt]{geometry}
\title{Preserving Accents in Real-Time Audio Translation: A Novel Approach for Linguistic Fidelity}
\author{Irfan Wani - Lovely Professional University}
\begin{document}    
\maketitle
    \begin{abstract}
    Our system takes the audio as input and translates it into the target language in real-time. The system is based on the concept of real-time audio translation. The system works in multiple steps, starting with taking the input from the user (primarily in english or detecting the language of the audio), converting the audio to text, taking the target language as the input from the user (like Hindi, Arabic etc) and then translating the text to the target language. Another step is the reverse of the previous steps where the text is converted back to audio and then played on the device.  NLP (Natural Language Processing) comes into play to make the whole thing working. NLP helps in translation, where it checks the grammer, words etc to make sure the translated text makes sense.
    \end{abstract}

\section{Introduction}
\section{Literature Review}
\section{Methodology}
\section{Results and Discussion}
\section{Future Scope}
\section{Conclusion}
\section{References}
\end{document}